---
title: "Seed morphology validation"
author: "James Hardwick"
date: "2024-12-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = 'C:/Users/jahar/Documents/Seed_transport_py/R_analysis/INNS_seed_transport')
```
# Introduction:
*Traitor* (Dayrell et al., 2023) was used to bulk process collected *Impatiens glandulifera* seeds for morphometric analysis. Each *Impatiens glandulifera* seed group was scanned on a high contrast surface and then morphology was extracted. Written in python using OpenCV and scikit-learn packages, each seed was extracted, aligned and measured. We validated *Traitor* based seed measurements with a manually measured 15 % subset (n ~ 180 seeds)using length and width as validation metrics. To quantify the degree of correlation between *Traitor* and manual seed width and length measurements Concordance Correlation Coefficient (CCC) was used - implemented with the *DescTools* library (Signorell, 2024). The code written below shows how CCC was calculated and provides validation plots. 

**Overall CCC values are:**
 - Length: 	*pc* = 0.57 (95 % CI [0.20 - 0.80]) 
 - Width:   *pc* = 0.88 (95 % CI [0.72 - 0.95])
 
 

# Libraries used:
```{r libraries, warning=FALSE}
library(tidyverse)
library(cowplot)
library(DescTools)
```


# data import and tyiding
```{r data import and tiyding}

morphology_validation_data <- read_csv("data/morphology_validation_data.csv")

```


# quantifiying validation 
```{r quantifying correaltion}
validation.length <- morphology_validation_data %>%
  select(group, manunal_length, traitor_length) %>%
  group_by(group) %>%
  summarise(mean_manunal_length = mean(manunal_length, na.rm = TRUE),
            mean_traitor_length = mean(traitor_length, na.rm = TRUE)) %>%
  filter(mean_traitor_length > 3)

# Assign variables
length.manunal <- validation.length$mean_manunal_length
length.traitor <- validation.length$mean_traitor_length

# length 
# Compute CCC with confidence intervals
length_ccc_result <- DescTools::CCC(
  x = length.manunal,
  y = length.traitor,
  ci = "z-transform",
  conf.level = 0.95
)

# Extract and round CCC and CI values
if (!is.null(length_ccc_result)) {
  length_pc <- round(as.numeric(length_ccc_result$rho.c), 2)  # CCC estimate
  length_ci_lower <- round(as.numeric(length_ccc_result$ci.lower), 2)  # Lower CI
  length_ci_upper <- round(as.numeric(length_ccc_result$ci.upper), 2)  # Upper CI
  
  # Combine into a CI string
  length_ci <- paste0("(", length_ci_lower, "-", length_ci_upper, ")")
  
  # Print result
  list(CCC = length_pc, Confidence_Interval = length_ci)
} else {
  stop("CCC computation failed: No result returned.")
}

# width 
# Summarize mean values for width
validation.width <- morphology_validation_data %>%
  select(group, manunal_width, traitor_width) %>%
  group_by(group) %>%
  summarise(mean_manunal_width = mean(manunal_width, na.rm = TRUE),
            mean_traitor_width = mean(traitor_width, na.rm = TRUE)) %>%
  filter(mean_traitor_width > 2)

# Assign variables
width.manunal <- validation.width$mean_manunal_width
width.traitor <- validation.width$mean_traitor_width
# Compute CCC with confidence intervals
width_ccc_result <- DescTools::CCC(
  x = width.manunal,
  y = width.traitor,
  ci = "z-transform",
  conf.level = 0.95
)

# Extract and round CCC and CI values
if (!is.null(width_ccc_result)) {
  width_pc <- round(as.numeric(width_ccc_result$rho.c), 2)  # CCC estimate
  width_ci_lower <- round(as.numeric(width_ccc_result$ci.lower), 2)  # Lower CI
  width_ci_upper <- round(as.numeric(width_ccc_result$ci.upper), 2)  # Upper CI
  
  # Combine into a CI string
  width_ci <- paste0("(", width_ci_lower, "-", width_ci_upper, ")")
  
  # Print result
  list(CCC = width_pc, Confidence_Interval = width_ci)
} else {
  stop("CCC computation failed: No result returned.")
}

```


# plotting validation data   
```{r validation plots}
 
  validation_length.plot <- ggplot(morphology_validation_data, aes(manunal_length, traitor_length))+
    geom_point()+
    geom_abline(intercept = 0, slope = 1, linetype = 'dashed')+
    ylim(0, NA)+
    xlim(0, NA)+
    xlab("Length Traitor (mm)")+
    ylab("Length Traitor (mm)")+
    annotate("text", x = 5, y = 1, label = "Pc 0.57 (95% CI)")+
    theme_bw()
  
  validation_width.plot <- ggplot(morphology_validation_data, aes(manunal_width, traitor_width))+
    geom_point()+
    geom_abline(intercept = 0, slope = 1, linetype = 'dashed')+
    ylim(0, NA)+
    xlim(0, NA)+
    xlab("Width Traitor (mm)")+
    ylab("Width Traitor (mm)")+
    annotate("text", x = 3.5, y = 0.5, label = "0.88 (95% CI)")+
    theme_bw()
  
  plot_grid(validation_length.plot, validation_width.plot, ncol = 1)

```





