---
title: "Impatiens glandulifera seed morphometrics"
author: "James Hardwick"
date: "2024-12-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = 'C:/Users/jahar/Documents/Seed_transport_py/R_analysis/INNS_seed_transport')
```


# Introduction:

This scripts takes morphmetric variable outputs from the *Traitor* python package (Dayrell et al., 2024) to calculate mean trait variables for Impatiens glandulifera seeds collected along the River Calder, West Yorkshire.Also, some additional morphometric variables, e.g. seed density and volume, are derived.  


## R Markdown

```{r libraries, warning=FALSE}
library(tidyverse)
library(cowplot)
library(RColorBrewer)
library(kableExtra)
```

# import data
```{r import and tidy data}
# seed morphology data 
seed_morphology_raw <- read_csv("data/seed_morphology_raw.csv")

seed_morphology <-seed_morphology_raw %>%
  group_by(group) %>%
  filter(length < 7.5) %>%
  summarise(mean_length_cm = mean(length)/10,
            mean_width_cm = mean(width)/10,
            mean_area_cm2 = mean(area)/10,
            mean_circularity = mean(circularity),
            mean_solidity = mean(solidity)) %>%
  mutate(group = tolower(group))

# seed mass data 
seed_mass <- read_csv("data/seed_mass.csv") %>%
  group_by(group)%>%
  summarise(mean_mass_g = mean(mass_g),
            mean_mass_mg = mean(mass_mg)) %>%
  mutate(group = tolower(group))

morph_mass <- seed_morphology %>%
  full_join(seed_mass, by = "group")

# seed settling velocity data 
seed_ws <- read_csv("data/seed_ws.csv", col_types = cols(...1 = col_skip())) 
  
seed_ws <- seed_ws %>%
  mutate(group = tolower(group))

```

# calcualte other morphometric variables and join with Ws
```{r morphometrics}

seed_morphometrics <- morph_mass %>%
 mutate(volume_cm3 = (4/3 * pi * (mean_width_cm/3)^3),
         density_gcm3 = mean_mass_g/volume_cm3,
         density_gcm3= if_else(density_gcm3 > 8, 
                                    density_gcm3 - (density_gcm3 * 0.95), density_gcm3))

```

```{r morph mass Ws}
ws_morph_mass <- morph_mass  %>%
  full_join(seed_ws, by ='group')
```


```{r summary table}
seed_ws_morphometrics <- read_csv("data/seed_ws_morphometrics.csv")

seed_ws_morphometrics <- read_csv("data/seed_ws_morphometrics.csv")
seed_morph_summary <- seed_ws_morphometrics %>%
  mutate(site_num = substr(group, 1, 2)) %>%
  group_by(site_num) %>%
  summarise(
    mean_width_cm = mean(mean_width_cm, na.rm = TRUE),
    sd_width_cm = ifelse(n() > 1, sd(mean_width_cm, na.rm = TRUE), NA),
    range_width_cm = diff(range(mean_width_cm, na.rm = TRUE)),
    
    mean_density = mean(density_gcm3, na.rm = TRUE),
    sd_density = ifelse(n() > 1, sd(density_gcm3, na.rm = TRUE), NA),
    range_density = diff(range(density_gcm3, na.rm = TRUE)),
    
    mean_ws_ms = mean(ws_ms.mean, na.rm = TRUE),
    sd_ws_ms = ifelse(n() > 1, sd(ws_ms.mean, na.rm = TRUE), NA),
    range_ws_ms = diff(range(ws_ms.mean, na.rm = TRUE))
  )

seed_morph_summary %>%
  kbl(caption = "Summary stats of seed traits") %>%
  kable_styling()


```

